---
title: "p8105_hw2_jae2186"
author: "jae2186 (Jennifer Estrada)"
date: "10/10/2021"
output:
  github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Problem 1

The following code chunk uploads the publicly available data of Mr. Trashwheel from May 2014 through June 2019 in an Excel spreadsheet format. 

The following changes have been made to the dataset:

* non-data entries are omitted (i.e. rows with notes/figures and columns containing notes),
* reasonable variable names have been applied, 
* rows that do not include dumpster-specific data are omitted, and
* the number of `sports_balls` is rounded to the nearest integer.


```{r datat_setup}

trash_wheel <-
  readxl::read_xlsx("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                    range = "A2:N408") %>% 
  drop_na() %>% 
  janitor::clean_names() %>% 
  rename(volume_cu_yds = volume_cubic_yards) %>% 
  mutate(sports_balls = round(sports_balls, 0)) 
 
```

The following code chunk reads, cleans, and combines precipitation data availble for 2018 and 2019.

```{r precip_data}

precip2019 <-
  readxl::read_xlsx("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                    range = "A2:B15",
                    sheet = 4)  %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  rename(precip_in = total) %>% 
  mutate (month = month.name[month], year = "2019") %>% 
  relocate(year)

precip2018 <-
  readxl::read_xlsx("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                    range = "A2:B15",
                    sheet = 5)  %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  rename(precip_in = total) %>% 
  mutate (month = month.name[month], year = "2018") %>% 
  relocate(year)

all_precip <- bind_rows(precip2018, precip2019)

```

In the dataset first presented describes the trash collected by Mr. Trashwheel from 2018 through 2019, taking in `r length(trash_wheel$dumpster)` dumpsters with an average of `r mean(trash_wheel$weight_tons)` tons of trash per dumpster, ulitmately collecting a  total of `r sum(trash_wheel$weight_tons)` tons of trash over the course of these 18 months.

Some quick statistics on the weather conditions and trash collected:

* <b>Precipitation</b>: Total precipitation in 2018 was `r sum(precip2018$precip_in)` inches. For the first half of 2019 precipitation was `r sum(precip2019$precip_in)` inches. A total of `r sum(all_precip$precip_in)` inches of precipitation fell over the 18-month observation period noted here.

* <b>Plastic bottles</b>: There were a over `r round(sum(trash_wheel$plastic_bottles)/10^6, 2)` million plastic bottles collected with a mean of `r round(mean(trash_wheel$plastic_bottles), 0)` plastic bottles per dumpster.

* <b>Glass bottles</b>: Comparatively there were only a total of `r sum(trash_wheel$glass_bottles)` glass bottles collected with a mean of `r round(mean(trash_wheel$glass_bottles), 2)` glass bottles per dumpster.

* <b> Sports balls</b>: Per dumpster there were an average of `r round(mean(trash_wheel$sports_balls), 2)` sports balls collected by Mr. Trashwheel during this time. When broken down by years:

  * In 2014 there were an average of `r round(mean(filter(trash_wheel, year == "2014")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2014")$sports_balls)` sports balls),
  * In 2015 there were an average of `r round(mean(filter(trash_wheel, year == "2015")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2015")$sports_balls)` sports balls),
  * In 2015 there were an average of `r round(mean(filter(trash_wheel, year == "2016")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2016")$sports_balls)` sports balls),
  * In 2015 there were an average of `r round(mean(filter(trash_wheel, year == "2017")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2017")$sports_balls)` sports balls),
  * In 2018 there were an average of `r round(mean(filter(trash_wheel, year == "2018")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2018")$sports_balls)` sports balls),
  * <b>And in 2019 there were an average of `r round(mean(filter(trash_wheel, year == "2019")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2019")$sports_balls)` sports balls).</b>
  
