---
title: "p8105_hw2_jae2186"
author: "jae2186 (Jennifer Estrada)"
date: "10/10/2021"
output:
  github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(DataCombine)
```

## Problem 1

The following code chunk uploads the publicly available data of Mr. Trashwheel from May 2014 through January 2021 in an Excel spreadsheet format. 

The following changes have been made to the dataset:

* non-data entries are omitted (i.e. rows with notes/figures and columns containing notes),
* reasonable variable names have been applied, 
* rows that do not include dumpster-specific data are omitted, and
* the number of `sports_balls` is rounded to the nearest integer.


```{r datat_setup}

trash_wheel <-
  readxl::read_xlsx("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                    range = "A2:N535") %>% 
  drop_na() %>% 
  janitor::clean_names() %>% 
  rename(volume_cu_yds = volume_cubic_yards) %>% 
  mutate(sports_balls = round(sports_balls, 0)) 
 
```

The following code chunk reads, cleans, and combines precipitation data availble for 2018 and 2019.

```{r precip_data}

precip2019 <-
  readxl::read_xlsx("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                    range = "A2:B14",
                    sheet = 6)  %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  rename(precip_in = total) %>% 
  mutate (month = month.name[month], year = "2019") %>% 
  relocate(year)

precip2018 <-
  readxl::read_xlsx("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                    range = "A2:B14",
                    sheet = 5)  %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  rename(precip_in = total) %>% 
  mutate (month = month.name[month], year = "2018") %>% 
  relocate(year)

all_precip <- bind_rows(precip2018, precip2019)

```

The dataset first presented describes the trash collected by Mr. Trashwheel from mid-May 2014 through January 2021, taking in `r length(trash_wheel$dumpster)` dumpsters with an average of `r mean(trash_wheel$weight_tons)` tons of trash per dumpster, ultimately collecting a  total of `r sum(trash_wheel$weight_tons)` tons of trash over the course of these years.

Some quick statistics on the weather conditions and trash collected:

* <b>Precipitation</b>: Total precipitation in 2018 was `r sum(precip2018$precip_in)` inches. For 2019 precipitation was `r sum(precip2019$precip_in)` inches. A total of `r sum(all_precip$precip_in)` inches of precipitation fell over the 2-year observation period noted here.

* <b>Plastic bottles</b>: There were a over `r round(sum(trash_wheel$plastic_bottles)/10^6, 2)` million plastic bottles collected with a mean of `r round(mean(trash_wheel$plastic_bottles), 0)` plastic bottles per dumpster.

* <b>Glass bottles</b>: Comparatively there were only a total of `r formatC(sum(trash_wheel$glass_bottles), format= "d", digits =2)` glass bottles collected with a mean of `r round(mean(trash_wheel$glass_bottles), 2)` glass bottles per dumpster.

* <b> Sports balls</b>: Per dumpster there were an average of `r round(mean(trash_wheel$sports_balls), 2)` sports balls collected by Mr. Trashwheel during this time. When broken down by years:

  * In 2014 there were an average of `r round(mean(filter(trash_wheel, year == "2014")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2014")$sports_balls)` sports balls),
  * In 2015 there were an average of `r round(mean(filter(trash_wheel, year == "2015")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2015")$sports_balls)` sports balls),
  * In 2015 there were an average of `r round(mean(filter(trash_wheel, year == "2016")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2016")$sports_balls)` sports balls),
  * In 2015 there were an average of `r round(mean(filter(trash_wheel, year == "2017")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2017")$sports_balls)` sports balls),
  * In 2018 there were an average of `r round(mean(filter(trash_wheel, year == "2018")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2018")$sports_balls)` sports balls),
  * <b>In 2019 there were an average of `r round(mean(filter(trash_wheel, year == "2019")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2019")$sports_balls)` sports balls),</b>
  * And in 2020 there were an average of `r round(mean(filter(trash_wheel, year == "2020")$sports_balls),2)`  sports balls collected (median of `r median(filter(trash_wheel, year == "2020")$sports_balls)` sports balls),
  


## Problem 2


Using the data from FiveThirtyEight looking at political party and economics measures, 

```{r pols_month, message=FALSE}
pols_month <-
  read_csv("./data/pols-month.csv") %>% 
  mutate(year = 0, month = 0, day = 0) %>% 
  separate(mon, c("year", "month", "day")) %>% 
  mutate(month = as.numeric(month), month = month.name[month], 
         year = as.numeric(year),
         prez = ifelse (prez_gop > 0, "gop", "dem")) %>% 
  select(-day, -prez_gop, -prez_dem) %>% 
  relocate(year, month, prez)
```


```{r snp, message=FALSE}

snp <- read_csv("./data/snp.csv") %>% 
  separate(date, c('month', 'day', 'year'), '/') %>% 
  select(-day) %>% 
  mutate(year = as.numeric(year), 
         month = as.numeric(month),
         month = month.name[month],
         year = ifelse(year < 50, year + 2000, year + 1900)) %>% 
  relocate(year, month)

```


```{r unemployment, message=FALSE}
unemploy <-
  read_csv("./data/unemployment.csv") %>% 
  rename(year = Year, 
         January = Jan, February = Feb, 
         March = Mar, April = Apr,
         May = May, June = Jun, 
         July = Jul, August = Aug, 
         September = Sep, October = Oct, 
         November = Nov, December = Dec) %>% 
  pivot_longer(January:December) %>%
  mutate(year = as.numeric(year)) %>%
  rename(month = name) %>% 
  arrange(year)
```


```{r joining, message=FALSE}
pol_unemploy <- 
  left_join(pols_month, snp) %>% 
  left_join(unemploy)

```

The first dataset entitled `pols_month` provided detailed information on the political party affiliation with regards to political leaders in various branches of government within the United States. The data provided the breakdown of political affiliation (i.e. either republican party noted as 'gop' or democratic party noted as 'dem') for the president as well as giving total counts for the number of senators (`sen_gop` or `sen_dem`), governors (`gov_gop` or `gov_dem`), and house representatives (`rep_gop` or `rep_dem`) affiliated with each party. The data provided monthly information on this political party breakdown, covering `r length(pols_month$year)` months, starting in `r pols_month$month[1]` `r min(pols_month$year)` and going until `r pols_month$month[length(pols_month$month)]` `r max(pols_month$year)`.

The second data set then provided dates and closing values for the Standard & Poor’s stock market index (S&P) with `r length(snp$year)` monthly values (taken on the 15th of each month) that started in `r snp$month[1]` `r min(snp$year)` and going until `r snp$month[length(snp$month)]` `r max(snp$year)`.

The third dataset provided the monthly rates of umemployment, covering `r length(unemploy$year)` months of data, starting in `r unemploy$month[1]` `r min(unemploy$year)` and going until `r unemploy$month[length(unemploy$month)]` `r max(unemploy$year)`.

## Problem 3

The following produces a table of the rank of popularity of the name "Olivia" over time as taken from the NYC Open data on the popularity of baby names. 


```{r cleaning_data, message=FALSE, warning=FALSE}
replace_eth <- 
  tibble(from = c("ASIAN AND PACI", 
                  "BLACK NON HISP",
                  "WHITE NON HISP"), 
         to = c("ASIAN AND PACIFIC ISLANDER", 
                "BLACK NON HISPANIC",
                "WHITE NON HISPANIC"))

baby_names <-
  read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct() %>% 
  mutate(ethnicity = replace(ethnicity, 
                             ethnicity == "ASIAN AND PACI", 
                             "ASIAN AND PACIFIC ISLANDER"),
         ethnicity = replace(ethnicity, 
                             ethnicity == "BLACK NON HISP", 
                             "BLACK NON HISPANIC"),
         ethnicity = replace(ethnicity, 
                             ethnicity == "WHITE NON HISP", 
                             "WHITE NON HISPANIC"))
```


```{r olivia}
olivia <-
  filter(baby_names, childs_first_name == "Olivia") %>% 
  select(-gender, -childs_first_name, -count) %>% 
  pivot_wider(names_from = year_of_birth, values_from = rank)
olivia
```


```{r male_no1}
male_no1 <-
  filter(baby_names, gender == "MALE" & rank == 1) %>% 
  select(-gender, - count, -rank) %>% 
    pivot_wider(names_from = year_of_birth, values_from = childs_first_name)

male_no1
```

```{r scatterplot}

wnh_names <-
  filter(baby_names, gender == "MALE" 
         & ethnicity == "WHITE NON HISPANIC" 
         & year_of_birth == 2016) %>% 
  select(-gender, -ethnicity, -year_of_birth)

ggplot(wnh_names, aes(rank, count)) + 
  geom_point() +
  labs(caption = "Figure 1. Count v. Rank for Names of Male, White Non-Hispanic Males Born in 2016 in NYC", x = "Name Rank", y = "Number of Children")
ggsave("p8105_hw2_p3_jae2186.pdf")

```


Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

